Exploring dimensions of biodiversity in Japanese ferns

### 2020-08-26 ms/SI.Rmd
Update MS

loadd(list = c(
  "japan_shp",
  "comm_scaled_list_0.1",
  "comm_scaled_list_0.2",
  "comm_scaled_list_0.3",
  "comm_scaled_list_0.4",
  "occ_point_data_ferns",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi"
),
cache = ja_fern_cache)

@@

>>>>

loadd(list = c(
  "japan_shp",
  "comm_scaled_list_0.1",
  "comm_scaled_list_0.2",
  "comm_scaled_list_0.3",
  "comm_scaled_list_0.4",
  "occ_point_data_ferns",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi",
  "phy_sig_results",
  "binary_sig_results"
),
cache = ja_fern_cache)

@@

```{r phy-sig}
phy_sig_results %>%
  rename(K = kval) %>%
  mutate(K = round(K, 3)) %>% 
  mutate(lambda = round(lambda, 3)) %>%
  mutate(lambda_pval = format(lambda_pval, digits = 3)) %>%
  write_csv(path = result_file(s_table("phy-sig"), "csv"))
```

**`r s_table("phy-sig")`**. `r s_table_cap("phy-sig")` K, Blomberg's *K*; lambda, Pagel's lambda.

```{r phy-sig-binary}
binary_sig_results %>%
  arrange(trait) %>%
  write_csv(path = result_file(s_table("phy-sig-binary"), "csv"))
```

**`r s_table("phy-sig-binary")`**. `r s_table_cap("phy-sig-binary")` num_present, number of times a trait was observed present; num_absent, number of times a trait was observed absent; D, Fritz and Purvis' *D*; prob_random, probability of obtaining observed trait distribution given random distribution of traits; prob_brownian, probability of obtaining observed trait distribution given random evolution by Brownian Motion. Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.

### 2020-08-26 ms/manuscript.Rmd
Update MS

loadd(list = c(
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns"
),
cache = ja_fern_cache)

@@

Our data include all ferns known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E). Japan consists of four main islands and thousands of smaller ones. The climate is subtropical in the south and highly seasonal in other parts of the country.

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (species, subspecies, and varieties). We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html). The final cleaned occurrence dataset included `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa). Hereafter, "taxa" refers to species, subspecies, and varieties unless otherwise indicated.

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of the cell) for defining communities (co-occurring species). At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell). We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. Continuous traits were measured on 10 randomly selected specimens per species. Fertile (i.e., spore-bearing) and sterile fronds were measured separately, then averaged. Qualitative traits were scored by observing voucher specimens. All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits. From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We compared SES of PD, RPD, FD, and RFD across bioregions using FIXME: add this when figure out spatial-corrected statistical method to use.

@@

We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit). We cropped grid cells to only include land area, then calculated the percent of protected area of each status type within grid cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness. For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. For taxonomic richness, cells ranked in the top >5% were considered highly diverse.

@@

@@

The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). The distribution of collection locations was highly skewed, with > 3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A). Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B). The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells. __ grid cells were excluded from the analysis due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.

@@

Grid cells with significantly low PD were almost all located on Honshu, the main island of Japan. A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). The area near Yakushima includes cells with both significantly high and low PD. Grid cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu. Grid cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.

@@

Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere. The vast majority of endemic grid cells are mixed, with very few paleo- or neoendemic cells. Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains). When we ran CANAPE with a reduced dataset including only taxa endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).

@@

Analysis of biogeographic regions using taxonomy (species names) resulted in the vast majority of grid cells classified into four regions: Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A). Using phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B). The other much smaller clusters are likely artifacts of insufficiently low taxon richness for accurate clustering in a small number of grid cells, and are not discussed further.

@@

The percentage of protected area in cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`). Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high status, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).

@@

### Conservation status and future outlook

Although the different facets of biodiversity have different distribution patterns, we find that grid cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric. While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups. Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change. One such other threat is herbivory by native Japanese deer, *Cervus nippon*. Populations of these deer have been increasing exponentially since the 1970s due to a combination of climate change and a decrease in hunting and natural predators. The deer are capable of decimating the herb layer when populations become dense and have been causing extensive damage even within national parks, which have the highest protection status. Although various efforts to prevent deer herbivory have been deployed such as exclusion fences, much work remains to be done to save vulnerable plant populations.

@@

>>>>

loadd(list = c(
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw"
),
cache = ja_fern_cache)

@@

Our data include all ferns known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E). Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). The climate is subtropical in the south and highly seasonal in other parts of the country. Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, and varieties unless otherwise indicated). We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. Continuous traits were measured on 10 randomly selected specimens per species. Fertile (\ie, spore-bearing) and sterile fronds were measured separately, then averaged. Qualitative traits were scored by observing voucher specimens. All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits. From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We compared SES of PD, RPD, FD, and RFD across bioregions using **FIXME: add this when figure out spatial-corrected statistical method to use**.

@@

We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit). We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness. For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. For taxonomic richness, cells ranked in the top >5% were considered highly diverse.

@@

# Calculate number of excluded cells due to low redundancy
n_excluded_cells <- nrow(comm_ferns_full) - nrow(comm_ferns)

@@

The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A). Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B). The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells. `r n_excluded_cells` grid-cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.

@@

Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan. A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). The area near Yakushima includes cells with both significantly high and low PD. Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu. Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.

@@

Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere. The vast majority of endemic grid-cells are mixed, with very few paleo- or neoendemic cells. Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains). When we ran CANAPE with a reduced dataset including only taxa endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).

@@

Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A). Phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B). The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`). Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).

@@

Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric. While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.

Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change. One such other threat is herbivory by native Japanese deer, *Cervus nippon*. Populations of these deer have been increasing exponentially since the 1970s due to a combination of climate change and a decrease in hunting and natural predators. The deer are capable of decimating the herb layer when populations become dense and have been causing extensive damage even within national parks, which have the highest protection status. Although various efforts to prevent deer herbivory have been deployed such as exclusion fences, much work remains to be done to save vulnerable plant populations.

### Limits of the study and future outlook

@@

```{r japan-map, fig.width = 5, fig.height = 5}
# Print map of Japan with labels of relevant places

# Set vectors of different areas to label
main_islands <- c("Hokkaido", "Shikoku", "Kyushu", "Honshu")
small_islands <- c("Okinawa", "Amamioshima", "Ogasawara", "Yakushima", "Izu Islands")
other <- c("Tokyo")

# Modify dataframe to include formatting for printing label on map
japan_points <- 
  japan_points_raw %>%
  as_tibble %>%
  mutate(
    label = case_when(
      # Print main islands in bold
      query %in% main_islands ~ glue::glue("**{query}**"),
      TRUE ~ query
    ),
    nudge = case_when(
      query == "Honshu" ~ "-0.5,-.7",
      query == "Tokyo" ~ "-1.2,0",
      query == "Shikoku" ~ "1,-.5",
      query == "Kyushu" ~ "1,-.6",
      query %in% small_islands ~ "2.5,0",
      TRUE ~ "0,0"
    )
  ) %>%
  separate(nudge, c("nudge_lon", "nudge_lat"), sep = ",") %>%
  mutate(across(c(nudge_lon, nudge_lat), as.numeric)) %>%
  mutate(
    lon_adj = lon + nudge_lon,
    lat_adj = lat + nudge_lat
  )

# Make map
ggplot(japan) + 
  geom_sf() +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  map_theme()

ggsave(result_file(figure("japan-map"), "pdf"), width = 5, height = 5, units = "in")
```

**`r figure("japan-map")`**. `r figure_cap("japan-map")`
